---
title: "R Notebook"
output: html_notebook
---

### 1. Loading Diamond Dataset
```{r}
install.packages('ggplot2')
library('ggplot2')
data(diamonds)
names(diamonds)
```

### 2. create a scatterplot of price vs x. using the ggplot syntax.
```{r}
ggplot(aes(x=x,y=price),data=diamonds)+
  geom_point()
```
### 3. What is the correlation between price and x,y,z? 
```{r}
with(diamonds,cor.test(price,x))
with(diamonds,cor.test(price,y))
with(diamonds,cor.test(price,z))
```

### 4.Create a simple scatter plot of price vs depth.
```{r}
ggplot(aes(x=depth,y=price),data=diamonds) +
  geom_point()
```

### 5.Change the code to make the transparency of the points to be 1/100 of what they are now and mark the x-axis every 2 units. See the instructor notes for two hints.
```{r}
ggplot(aes(x=depth,y=price),data=diamonds) +
  geom_point(alpha=1/100)+
  scale_x_continuous(breaks=seq(0,80,2))
```
### 6. Based on the scatterplot of depth vs price, most diamonds are between what values of depth?
Ans. 60 to 64

### 7. What is the correlation of depth vs price?
```{r}
with(diamonds,cor.test(depth,price))
```

### 8. Create a scatterplot of price vs carat and omit the top 1% of price and carat values.
```{r}
ggplot(aes(x=carat,y=price),data=diamonds)+
  geom_point()
```

