---
title: "R Notebook"
output: html_notebook
---

### 1. Loading Diamond Dataset
```{r}
install.packages('ggplot2')
library('ggplot2')
data(diamonds)
names(diamonds)
```

### 2. create a scatterplot of price vs x. using the ggplot syntax.
```{r}
ggplot(aes(x=x,y=price),data=diamonds)+
  geom_point()
```
### 3. What is the correlation between price and x,y,z? 
```{r}
with(diamonds,cor.test(price,x))
with(diamonds,cor.test(price,y))
with(diamonds,cor.test(price,z))
```

### 4.Create a simple scatter plot of price vs depth.
```{r}
ggplot(aes(x=depth,y=price),data=diamonds) +
  geom_point()
```

### 5.Change the code to make the transparency of the points to be 1/100 of what they are now and mark the x-axis every 2 units. See the instructor notes for two hints.
```{r}
ggplot(aes(x=depth,y=price),data=diamonds) +
  geom_point(alpha=1/100)+
  scale_x_continuous(breaks=seq(0,80,2))
```
### 6. Based on the scatterplot of depth vs price, most diamonds are between what values of depth?
Ans. 60 to 64

### 7. What is the correlation of depth vs price? Basedd on the correlation coefficient would you depth to predict the price of diamond?
```{r}
with(diamonds,cor.test(depth,price))
```
As the correlation coefficient value seems to very low, we can say that depth and price doesn't have a strong relationship, So we cannot use the depth to predict the price of a diamond.  


### 8. Create a scatterplot of price vs carat and omit the top 1% of price and carat values.
```{r}
ggplot(aes(x=carat,y=price),data=diamonds)+
  geom_point()+
  xlim(0,quantile(diamonds$carat,0.99))+
  ylim(0,quantile(diamonds$price,0.99))+
  geom_smooth()
```

### 9.Create a scatterplot of price vs. volume (x  y  z). This is a very rough approximation for a diamond's volume.
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z
ggplot(aes(x=volume,y=price),data=diamonds)+
  geom_point()
```

### 10.How many diamonds have zero volume?
```{r}
sum(diamonds$volume==0)
```

### 11.What is the correlation of price and volume? Exclude diamonds that have volume of zero or that are greater than or equal to 800.
```{r}
with(subset(diamonds,volume > 0 & volume <= 800),cor.test(price,volume))
```

### 12.Subset the data to exclude diamonds with a volume greater than or equal to 800. Also, exclude diamonds with a volume of 0. Adjust the transparency of the points and add a linear model to the plot.
```{r}
ggplot(aes(x=volume,y=price),data=subset(diamonds,volume > 0 & volume <= 800))+
  geom_point(alpha=1/5)+
  geom_smooth(method = 'lm')
```

### 13.Use the function dplyr package to create a new data frame containing info on diamonds by clarity.
```{r}
library(dplyr)

diamonds_by_clarity <- group_by(diamonds,clarity)

diamondsByClarity <- summarise(diamonds_by_clarity,
                               mean_price = mean(price),
                               median_price = median(price),
                               min_price = min(price),
                               max_price = max(price),
                               n = n())


```

### 14. create two bar plots on one output image using the grid.arrange() function from the package gridExtra.
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

```

```{r}
library(gridExtra)
p1 <- ggplot(data=diamonds_mp_by_clarity,aes(x=clarity,y=mean_price))+
  geom_bar(stat = "identity")

p2 <- ggplot(data=diamonds_mp_by_color,aes(x=color,y=mean_price))+
  geom_bar(stat = "identity")

p3 <- ggplot(data=diamonds_mp_by_cut,aes(x=cut,y=mean_price))+
  geom_bar(stat = "identity")

grid.arrange(p1,p2,p3)
```

