White Wine Data Exploration
========================================================
This tidy data set contains 4,898 white wines with 11 variables on quantifying the chemical properties of each wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

This Report explores the White Wine dataset to findout which chemical properties influence the quality of wines.

### 1. Required imports and libraries.
```{r}
library(ggplot2)
```

### 2. Loading the dataset.
```{r}
setwd('E:/git/Udacity-Data-Analyst-Nano-Degree/P4 Explore And Summarize Data/EDA_PROJECT_WINE_DATA_EXPLORATION')
wines <- read.csv('wineQualityWhites.csv')
head(wines)
```

### 3. Dimensions of the dataset.
```{r, echo=TRUE}
dim(wines)
```

### 4. Description of the variables and thier types.
```{r}
str(wines)
```

### 5. Summaries of numerical variables present in dataset. 
```{r}
summary(wines)
```

Now lets explore the distribution of fixed acidity in our samples.The fixed.acidity peaks at 6.5, but most of the samples seems to have acidity between 6 to 7.5 .
```{r}
ggplot(aes(x=fixed.acidity),data=wines)+
  geom_histogram(binwidth = 0.25,color="black")+
  scale_x_continuous(breaks = seq(3,15,0.5))
```


Now lets explore the distribution of volatile.acidity in our dataset.
```{r}
ggplot(aes(x=volatile.acidity),data=wines)+
  geom_histogram(binwidth =0.01,color='black')+
  scale_x_continuous(breaks = seq(0,1.2,0.05))
```

Now lets explore the contents of citric.acid in our dataset.
```{r}
ggplot(aes(x=citric.acid),data=wines)+
  geom_histogram(binwidth = 0.05,color='black')+
    scale_x_continuous(breaks = seq(0,1.7,0.1))
```

Now Lets explore the distribution of residual.sugar in our sample.
```{r}
ggplot(aes(x=residual.sugar),data=wines)+
  geom_histogram(color="black",binwidth = 1)
```

The above distribution seems to be skewed and long tailed due to some outliers. Lets plot the data using log transformation to see the more detailed distribution. 
```{r}
ggplot(aes(x=residual.sugar),data=subset(wines))+
  geom_histogram(binwidth = 0.05,color='black')+
  scale_x_log10(breaks=c(1,3,5,7,10,3,15,20,25,31,65))
```

Now lets explore the distribution of chlorides in our dataset.
```{r}
ggplot(aes(x=chlorides),data=wines)+
  geom_histogram(binwidth=0.01,color="black")
```

As most of the values in the above distribution lie between 0.0 to 0.1 and the graph seems to be a bit long tailed towards right, lets use log scale to get more disbursed distribution.
```{r}
ggplot(aes(x=chlorides),data=wines)+
  geom_histogram(binwidth =0.05,color='black')+
  scale_x_log10(breaks=seq(0.01,0.35,0.05))
```

Now Lets see the distribution free.sulfur.dioxide in ou dataset.
```{r}
ggplot(aes(x=free.sulfur.dioxide),data=wines)+
  geom_histogram(binwidth=10,color='black')
```

Now Lets see the distribution of total.sulfur.dioxide in our dataset.
```{r}
ggplot(aes(x=total.sulfur.dioxide),data=wines)+
  geom_histogram(binwidth = 5,color="black")
```

```{r}
ggplot(aes(x=density),data=wines)+
  geom_histogram(binwidth = 0.001,color="black")+
  xlim(0.985,1.04)
```

```{r}
ggplot(aes(x=pH),data=wines)+
  geom_histogram(binwidth = 0.03,color="black")
```

```{r}
ggplot(aes(x=sulphates),data=wines)+
  geom_histogram(binwidth=0.02,color="black")
```

```{r}
ggplot(aes(x=alcohol),data=wines)+
  geom_histogram(binwidth = 0.1,color="black")
```

```{r}
ggplot(aes(x=quality),data=wines)+
  geom_histogram(bins = 7,color="black")+
  scale_x_continuous(breaks = seq(3,9,1))
```
