White Wine Data Exploration
========================================================
This tidy data set contains 4,898 white wines with 11 variables on quantifying the chemical properties of each wine. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent).

This Report explores the White Wine dataset to findout which chemical properties influence the quality of wines.

## Univariate Plot Section

### 1. Required imports and libraries.
```{r message=FALSE, warning=FALSE}
library(ggplot2)
```

### 2. Loading the dataset.
```{r}
setwd('E:/git/Udacity-Data-Analyst-Nano-Degree/P4 Explore And Summarize Data/EDA_PROJECT_WINE_DATA_EXPLORATION')
wines <- read.csv('wineQualityWhites.csv')
head(wines,3)
```

### 3. Dimensions of the dataset.
```{r, echo=TRUE}
dim(wines)
```

### 4. Description of the variables and thier types.
```{r}
str(wines)
```

### 5. Summaries of numerical variables present in dataset. 
```{r}
summary(wines)
```

Now lets explore the distribution of fixed acidity in our samples.The fixed.acidity peaks at 6.5, but most of the samples seems to have acidity between 6 to 7.5 .
```{r}
ggplot(aes(x=fixed.acidity),data=wines)+
  geom_histogram(binwidth = 0.25,color="black")+
  scale_x_continuous(breaks = seq(3,15,0.5))
```

Now lets explore the distribution of volatile.acidity in our dataset. High amounts of this content(acetic acid) can lead to unpleasant, vinegar taste to wines which will eventually effect the quality of wines.
```{r}
ggplot(aes(x=volatile.acidity),data=wines)+
  geom_histogram(binwidth =0.01,color='black')+
  scale_x_continuous(breaks = seq(0,1.2,0.05))
```

Now lets explore the contents of citric.acid in our dataset. Citric acid found in samll quantities, can add freshness and flavour to wines.
```{r}
ggplot(aes(x=citric.acid),data=wines)+
  geom_histogram(binwidth = 0.05,color='black')+
    scale_x_continuous(breaks = seq(0,1.7,0.1))
```

Now Lets explore the distribution of residual.sugar in our sample.
```{r}
ggplot(aes(x=residual.sugar),data=wines)+
  geom_histogram(color="black",binwidth = 1)
```

The above distribution seems to be skewed and long tailed due to some outliers. Lets plot the data using log transformation to see the more detailed distribution. 
```{r}
ggplot(aes(x=residual.sugar),data=subset(wines))+
  geom_histogram(binwidth = 0.05,color='black')+
  scale_x_log10(breaks=c(1,3,5,7,10,3,15,20,25,31,65))
```

Transformed the long-tailed data to better understand the distribution of residual.sugar. The transformed residual.sugar appears to be bimodal with two peaks. As mentioned in dataset descripton that wines which have residual.sugar content above 45 are considered to be sweet, lets find out how many of them are there.
```{r}
table(wines$residual.sugar > 45)
```



Now lets explore the distribution of chlorides in our dataset.
```{r}
ggplot(aes(x=chlorides),data=wines)+
  geom_histogram(binwidth=0.01,color="black")
```

As most of the values in the above distribution lie between 0.0 to 0.1 and the graph seems to be a bit long tailed towards right, lets use log scale to get more disbursed distribution.
```{r}
ggplot(aes(x=chlorides),data=wines)+
  geom_histogram(binwidth =0.05,color='black')+
  scale_x_log10(breaks=seq(0.01,0.35,0.05))
```

Now Lets see the distribution free.sulfur.dioxide in ou dataset.
```{r}
ggplot(aes(x=free.sulfur.dioxide),data=wines)+
  geom_histogram(binwidth=5,color='black')
```

The above plot shows that most of the data lies below 100, so lets zoom into the plot by excluding some outliers.
```{r message=FALSE, warning=FALSE}
ggplot(aes(x=free.sulfur.dioxide),data=wines)+
  geom_histogram(binwidth=2,color='black')+
  scale_x_continuous(breaks=seq(0,100,5),limits = c(0,100))
```

Now Lets see the distribution of total.sulfur.dioxide in our dataset.
```{r}
ggplot(aes(x=total.sulfur.dioxide),data=wines)+
  geom_histogram(binwidth = 5,color="black")
```

In the above graph most of the data lies below 250, Lets zoom into the plot by excluding some outliers.
```{r message=FALSE, warning=FALSE}
ggplot(aes(x=total.sulfur.dioxide),data=wines)+
  geom_histogram(binwidth = 5,color="black")+
  scale_x_continuous(limits = c(0,250),breaks=seq(0,250,10))
```


Now lets explore the density of wines in our dataset. The desity of wine is close to the density of water but it also depends upon the percent alcohol and sugar content.
```{r}
ggplot(aes(x=density),data=wines)+
  geom_histogram(binwidth = 0.001,color="black")
```

As we know that density of water is one and density of wine is close to that of water, we see that most of the points lie below one unit. Lets zoom into the plot to see more disbursed distribution. 
```{r message=FALSE, warning=FALSE}
ggplot(aes(x=density),data=wines)+
  geom_histogram(binwidth = 0.0002,color="black")+
  xlim(0.9871,1.0001)
```

Now lets explore the pH levels of the wines in our dataset.pH describes how acidic or basic a wine is on a scale from 0 (very acidic) to 14 (very basic); most wines are between 3-4 on the pH scale.
```{r}
ggplot(aes(x=pH),data=wines)+
  geom_histogram(binwidth = 0.05,color="black")+
  scale_x_continuous(breaks=seq(2.7,3.9,0.1))
```

Now lets explore the distribution of sulphates in our dataset. Sulphates are a wine additive which can contribute to sulfur dioxide gas (S02) levels, wich acts as an antimicrobial and antioxidant
```{r}
ggplot(aes(x=sulphates),data=wines)+
  geom_histogram(binwidth=0.02,color="black")
```

Now lets explore the distribution amounts of alcohol in our dataset.
```{r}
ggplot(aes(x=alcohol),data=wines)+
  geom_histogram(binwidth = 0.1,color="black")
```

Now lets see explore the distribution of quality. The variable Quality appears to be a int variable lets transform this variable to factor variable, so that it can be useful later in plotting categorical graphs.
```{r}
table(wines$quality)
```

```{r}
ggplot(aes(x=quality),data=wines)+
  geom_histogram(bins = 7,color="black")+
  scale_x_continuous(breaks = seq(3,9,1))
```

## Univariate Analysis

### What is the structure of your dataset?
There are 4,898 diamonds in the dataset with 11 features(fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides,     free.sulfur.dioxide, total.sulfur.dioxide, density, pH, sulphates, alcohol) and quality being the sensory data.

Observations:

1. Residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide have huge outliers, which are too far away from the mean.
2. There is only one wine in the dataset which can be considered as sweet.
3. Most of the wines in the dataset have the quality as 5 or 6, only few wines are of the best quality(8 0r 9).

### What is/are the main feature(s) of interest in your dataset?
The main feature of interest is "Quality". I would like to determine what all factors are affecting the quality of wine.

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?
Features like contents of acids,amounts of sulfur and chlorides, amount of residual.sugar and percent of alcohol might help in investigation.  

### Did you create any new variables from existing variables in the dataset?
No I didn't create any new varaiable, but transformed the existing variable "Quality" to a factor variable, So that it can be later useful to plot categorical graphs in terms of quality.

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

Of the features investigated, residual.sugar and chlorides have unusal distributions. I log transformed these skewed distributions, so that we can get a clear and more disbured distribution of data.

The transformed distribution for residual.sugar appears to be bimodal with peaks around 1g/dm^3 and 7g/dm^3. The Chlorides distribution is heavly populated between 0.0 to 0.1, so I used log transfromation, which transfroms the data on a ten-fold sacle, so that we can get more widely populted distribution of chlorides.     


## Bivariate Plot Section

In this section, I will be exploring the two variable relationships.
```{r}
cor(wines)
```

As our main feature of interest lies in "Quality", Lets see how quality is getting effected by other variables like alcohol. Quality and alcohol seem to have a bit strong relationship because they have positive correlation co-efficent of 0.435.
```{r}
ggplot(aes(x=factor(quality),y=alcohol),data=wines)+
  geom_boxplot()
```








