---
title: "R Notebook"
output:
  html_document: default
---

### 1. Laoding Diamond Dataset.
```{r}
#install.packages('ggplot2')
library('ggplot2')
data(diamonds)
```

### 2. How many observations are in the dataset.
```{r}
data(diamonds)
nrow(diamonds)
```
### 3. How many variables are there in the dataset ?
```{r}
ncol(diamonds)
```
### 4. How many ordered factors are there in the dataset?
```{r}
str(diamonds)
```
### 5. What leter represents the best color for a diamonds?
```{r}
levels(diamonds$color)
```

### 6. Create Histogram for price data.
```{r}
ggplot(aes(x=price),data=diamonds)+
  geom_histogram(fill='#F79420',color='black')
```
### 7. Find summary statistics like mean and median.
```{r}
mean(diamonds$price)
median(diamonds$price)
```
### 8. How many diamonds cost less than 500?.
```{r}
sum(diamonds$price < 500)
```
### 9. How many diamonds cost less than 250?.
```{r}
sum(diamonds$price < 250)
```
### 10. How many diamonds cost 15000 or more?.
```{r}
sum(diamonds$price >= 15000)
```
### 11. Explore the largest peak in price histogram.
```{r}
ggplot(aes(x=price),data = diamonds)+
  geom_histogram(color='black',fill='#F79420',binwidth = 50)+
  scale_x_continuous(limits = c(200,1000),breaks = seq(300,950,50))
```
### 12. Break out the histogram prices by cut.
```{r}
ggplot(data=diamonds,aes(x=diamonds$price))+
  geom_histogram(color='black')+
  facet_wrap(~cut)
```
### 13. Summary stattistics to find min and max by cut.

```{r}
by(diamonds$price,diamonds$cut,summary)
```

```{r}
ggplot(data=diamonds,aes(x=diamonds$price))+
  geom_histogram(color='black',fill='#F79420')+
  facet_wrap(~cut,scales = 'free_y')
```

### 14.Create a histogram of price per carat and facet it by cut.
```{r}
ggplot(aes(x=price/carat),data=diamonds)+
  geom_histogram(color='black',fill='#F79420')+
  scale_x_log10()+
  facet_wrap(~cut)
```

### 15. Investigate the price of diamonds using boxplots.
```{r}
ggplot(aes(x=color,y=price),data=diamonds)+
  geom_boxplot(fill='#F79420')

by(diamonds$price,diamonds$color,summary)
  
```

### 16.Investigate the price per carat of diamonds across the different colors of diamonds using boxplots.
```{r}
ggplot(aes(y=price/carat,x=color),data=diamonds)+
  geom_boxplot(fill='#F79420')
```

### 17. Investigate the weights of diamonds using a frequency ploygon.
```{r}
ggplot(aes(x=carat),data=diamonds)+
  geom_freqpoly(binwidth=0.1,color='orange')+
  scale_x_continuous(limits = c(0,2),breaks=seq(0,2,0.1))
```

