---
title: "Facebook Friends Birthday Data Analysis"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
  word_document: default
---

## 1. Loading the data file and required packages
```{r warning=FALSE}
#install.packages('lubridate')
library(lubridate)
fb <- read.csv('birthdays.csv')
```

## 2. Lets take look at the dataset using the head command.
```{r}
head(fb)
```

## 3. Lets remove the columns End and duration, as they might not be much useful for exploration.
```{r}
fb <- subset(fb,select = -c(End,Duration))
head(fb)
```

## 4. Lets convert the Start column data to date format.
```{r}
fb$Start <- as.Date(fb$Start,format='%m/%d/%Y')
head(fb$Start)
```

## 5. Lets add month column to dataset.
```{r}
fb$month <- format(fb$Start,'%b')
head(fb)
```

# Now Lets Explore The Dataset.

## 7. How many birthdays are in each month?
```{r}
table(fb$month)
```

## 8. How many people share My birthday?
```{r}
myBdayDate = as.Date('10/12/2017','%m/%d/%Y')
shareMyBday <- subset(fb,fb$Start == myBdayDate)
shareMyBday
```

## 9. Which month contains the most number of birthdays?
```{r}
names(which(table(fb$month) == max(table(fb$month))))
```

## 10. Plot a bar chart to show number of birthdays in each month
```{r warning=FALSE}
#install.packages("ggplot2")
library(ggplot2)
```



```{r}
fb$month <- factor(fb$month,ordered = TRUE,levels = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
ggplot(data=fb,aes(x=month))+
  geom_bar(color='black',fill='#F79420',width = 0.6)+
  geom_text(stat="count",aes(label=..count..),vjust=-0.8)+
  scale_y_continuous(limits = c(0,55),breaks = seq(0,55,10))+
  labs(x="Different  Months",y="No of Birthdays",title="Month Wise facebook Birthdays Distribution.")+
  theme_bw()+
  theme(axis.title = element_text(size=14),plot.title =element_text(hjust=0.5,size =14,face="bold"))
```


